---
title: "p8105_hw2_js5959"
author: "Junzhe Shao"
date: "10/9/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(readxl)
library(tidyverse)
library(lubridate)
library(ggrepel)
```


## Problem 1
### read excel sheet "Mr. Trash Wheel" and do the data cleaning
```{r warning = FALSE}
#read excel sheet "Mr. Trash Wheel" and do the data cleaning
mr_tw_dt = read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel", skip = 1) %>% 
  janitor::clean_names() %>% 
  select(-c(x15,x16,x17)) %>% # remove extra columns
  filter(dumpster != "Grand Total") %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls)
  )
#display first ten rows 
kable(mr_tw_dt[1:10,])
```
### clean precipitation for 2018 and 2019
```{r warning = FALSE}
# clean precipitation for 2018 and 2019
pr_19_dt = read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2019 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(
    month = month.name[month],
    year = 2019)
pr_18_dt = read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2018 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(
    month = month.name[month],
    year = 2018)
pr_dt = bind_rows(pr_19_dt,pr_18_dt)
kable(pr_dt)
```
### A description of both resulting datasets:
observation of cleaned Mr. Trash Wheel data is 453, variables in that dataset indicating the a summary of date, total amount of trash collected and their specific types.
observation of combined 2018 and 2019 Precipitation data is 24 month. The total Precipitation of 2018 is `r pr_dt %>% filter(year == 2018) %>% pull(total) %>% sum()`
the medium number of sports balls in a dumpster in 2019 is `r mr_tw_dt %>% filter(year == 2019) %>% pull(sports_balls) %>% median()`

## Problem 2
### clean the data in pols-month.csv.
```{r warning = FALSE}
pols_month_dt = read_csv("./fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon,c("year","month","day")) %>% 
  mutate(
    month = month.name[as.numeric(month)],
    president = ifelse(prez_gop == 0, "dem","gop")# replace 0 with dem 
  ) %>% 
  select(-c(prez_dem,prez_gop,day))#remove columns
#display first ten rows
kable(pols_month_dt[1:10,])
```
### clean snp.csv 
```{r warning = FALSE}
snp_dt = read_csv("./fivethirtyeight_datasets/snp.csv") %>% 
  mutate(
    date = mdy(date)
  ) %>% 
  separate(date,c("year","month","day")) %>% 
  mutate(
    month = month.name[as.numeric(month)],
    year = as.numeric(year),
    year = ifelse(year > 2021,year-100,year),
    year = as.character(year)
  ) %>% 
  select(-day)
kable(snp_dt[1:10,])    
```
### clean unemployment data
```{r}
unemp_dt = read_csv("./fivethirtyeight_datasets/unemployment.csv") %>% 
    pivot_longer(
      Jan:Dec,
      names_to = "month", 
      values_to = "percentage_unemployment") %>% 
    mutate(date = str_c(Year, month, sep = "-"),
           date= ym(date)) %>% 
  separate(date, c("year", "month","day")) %>% 
  mutate(
    month = month.name[as.numeric(month)]
  ) %>% 
  select(-c(day,Year))
kable(unemp_dt[1:10,])
```
```{r}
res_dt <- left_join(pols_month_dt, snp_dt,by =c("year"="year","month"="month"),na_matches="never")
res_dt <- left_join(res_dt, unemp_dt,by =c("year"="year","month"="month"),na_matches="never") %>% 
  arrange(year,month)# arrange by order of date
kable(res_dt[1:10,])
```
### A bref discription

The finial res_dt dataset includes all three parts above.There are indicator of whether the president was republican/democratic, the number of republican/democratic governors/representatives/senators,closing values of the S&P stock index and the unemployment rate and variables indicating date.  The final res data set has `r nrow(res_dt)`entries. The year spanning over from `r min(pull(res_dt, year))` to `r max(pull(res_dt, year))`.

## Problem3

### A basic data cleaning with notes in questions
```{r echo = T, results = 'hide'}
#According to the note, we need to specify the ethnicity name and control the case of names.

pop_names_dt = read_csv("Popular_Baby_Names.csv") %>% 
  janitor::clean_names()%>% 
    separate(ethnicity, "ethinic", extra = "drop") %>% # ethinic dupicated remove
  mutate(
    childs_first_name = tolower(childs_first_name)# lower case
  ) %>% distinct()# remove duplicated rows
  
```
There are `r nrow(pop_names_dt)` entries after remove duplicated rows.

### female named with olivia
```{r}
olivia_female_dt = pop_names_dt %>% 
  filter(childs_first_name == "olivia",gender == "FEMALE") %>% 
  select(-c(childs_first_name,gender,count)) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank")
kable(olivia_female_dt)
```
### The most popluar male man trending for all years
```{r}
pop_male_name = pop_names_dt %>% 
  filter(rank == 1,gender == "MALE") %>% 
  select(-c(rank,gender,count)) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  )
kable(pop_male_name)
```
### scatter plot for white male in 2016:
```{r}
male_white_2016 = pop_names_dt %>% 
  filter(gender == "MALE",ethinic == "WHITE",year_of_birth == 2016) 

ggplot(male_white_2016, aes(x = rank, y = count)) +geom_point(colour = "skyblue") +geom_text_repel(label = male_white_2016$childs_first_name)# add not overlapping labels
```



